Sketch Recognition using Particle Swarm Optimizations

### 2009-04-16 trunk/paper2/Paper2.tex
after the my edit of introduction and review

\documentclass{article}
\usepackage{spconf}
\usepackage{amsmath}
\usepackage[dvips,pdftex]{graphicx}
\usepackage{subfigure}
\usepackage{named}
\usepackage[dvips,pdftex]{hyperref}
\title{Sketch Recognition using Particle Swarm Optimizations}
\twoauthors
  {Maha El Meseery, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
	{Signals Processing Group \\ Computers and Systems Department \\
  Electronic Research Institute\\Cairo, Egypt\\
	melmeseery@eri.sci.eg, samia@eri.sci.eg}
  {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
	{ Computer Engineering Department\\
	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
	magdafayek@gmail.com,ndarwish@ieee.org}
\begin{document}
\maketitle
\begin{abstract}
Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. The results show that using PSO improves segmentation results which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.  
\end{abstract}
\section{Introduction}
The field of symbol recognition has gained interest in the last few years. Scientists generally and engineers specifically express thoughts and designs using sketches. Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking. Design engineers need a powerful computer based system with symbol recognition capabilities to help them design, manipulate and store sketches more effectively than using only papers. Increasing the interaction between computers and users in sketch and CAD systems has been the main reason for the emerging of a few advanced sketch recognition systems.

Sketch recognition is divided into three main steps a) preprocessing, b) segmentation and c) symbol recognition. The preprocessing phase captures user input stroke points and collects basic information about the stroke. This is followed by proceeding to remove noise and computing basic statistical and geometrical information. In the segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments. In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier.

This paper introduces a new sketch recognition system, which uses particle swarm optimization (PSO) to optimally segment user's strokes. Users can draw symbols using any number of strokes in any order they like. The system segments each stroke using two different PSO algorithms. The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}. Our algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results. To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs. After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced. The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed. % uses the computed feature vector to classify the segments into one of the previously trained classes.

 The remaining of the paper is as follows; the literature review is in Section \ref{sec:review}. Section \ref{sec:ParticleSwarmAlgorithm} introduces Particle Swarm Optimization. Section \ref{Sysdisc} describes the system in details.  Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively. Section \ref{sec:Experiments} describes the experiments that were preformed. We conclude in Section \ref{ConclusionandFutureWork}.

\section{Literature Review}
\label{sec:review}
A wide variety of techniques were suggested for sketch recognition.  A hybrid algorithm was introduced in \cite{earlyprocess} where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set. However, this system is limited to recognizing only specific simple geometric shapes with a set of low level recognizers. Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from input stroke.  Yu \cite{meanshift10} introduced a \textit{feature area} for each primitive and then computed the segmentation error for different types of primitives based on the computed \textit{feature area}. His system achieved good accuracy in simple shapes (square, ellipse,...etc) but did not perform well in complex shapes. Paulson and Hammond \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%   accuracy. However, their system similar to all low level recognizers in the way that it identifies a small set of simple shapes.% A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.

A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. Chen et al. \cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system. Also, they did not take advantage of the curvature or local geometric properties of the digital curve. Yin \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts  Yin \cite{PolygonApproximationPSO} method and tries to improve it by adding curvature and other local information while segmenting strokes to achieve better segmentations. The next section presents the general particle swarm algorithm which is used in both our segmentation algorithms. % are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}

@@

After the system computes the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvature. However, using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of each of the four calcuated information  is calculated. This mean is taken as the threshold \textit{th} which is used to separate the curve into regions ($R_i$) (Figure\ref{fig:timediff}); each region $R_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}. Those regions are further processed to find the maximum point $Max(R_i)$ of each region $R_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. The system repeats this process for curvature, time difference and speed information. All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate point $P_{pd}$ as the preprocessing. It is clear that some of the $P_{pd}$ points are redundant as specified in the preprocessing stage. % (as shown are redundant)
\begin{figure}
	\centering
		\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{images/ppd.jpg}}
		\hfill
			\subfigure[DPSO algorithm encoding]{	\label{fig:LabelsPSO} \includegraphics[scale=0.7]{images/pso1.jpg}}			
	\caption{Segmentation}% a) Possible dominate point b) Particle encoding  
	\label{fig:pso1}
\end{figure}

\subsection{Segmentation}
\label{seg}
In the segmentation stage the input stroke is divided into a set of primitives. As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}. The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two furthur segmentation algorithms that divide the stroke to either lines or lines and curves. The next section describes in details the ellipse detection algorithm and the two PSO algorithms used to divide the stroke. % The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing. The output is a set of dominant points which are connected by either lines or curves.

%///AllBlockDiagram.jpg  you can change the block to the main block. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.5]{images/blockSmall.jpg}
	\caption{Segmentation Block}
	\label{fig:segblock}
\end{figure}
\subsection{Ellipse fitting} 
%The process starts by computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the ellipse center. The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.  
t
This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the center of the ellipse. The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box. The least square fitting algorithm \cite{chernov-2003} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})

@@

 where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke. A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.  After few iteration, the final fit error of the estimated ellipse is reported. Another measure is used to compute the efficiency of the final estimated ellipse. Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered. This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fittted as a pratial or full ellipse.

@@

 \subsubsection{\textit{Hybird Fitting \textsl{AlgS2}}}
The algorithm has the same problem formulation but different fitness and error functions are used. An attempt is made to fit each segment into a line or circular arc. The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$. The approximation with the lower error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}. The sum of the approximation errors of all segments is the total error of the particle.  The total error of the particle is computed by equation %\ref{eq:errorSwarm2}
 \begin{equation}
E=\sum\nolimits_{i = 0}^M e(D_i) 
\label{eq:errorSwarm2}
\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.  The fitness is computed by the equation %\ref{eq:fitnessSwarm2} 
\begin{equation}
\max fitness(P_i ) = \frac{1}{{E \times M^k }}
\label{eq:fitnessSwarm2}
\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments. The larger $k$ is, the more effect the number of segments will have. For our system, $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.

@@

\section{Experimental}
\label{sec:Experiments}
The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system. The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. The dataset was divided into training set and test set. Four different splits for the training and test data are generated from the dataset. The results displayed are the average recognition accuracy of the four splits. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test. %In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.  
\begin{figure}
\centering 
\label{fig:symbolSet}
\fbox{ \parbox{5cm}{% 
		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
		\caption{The Symbol Set} Symbols used in the dataset.
\end{figure}
\subsection{Algorithms comparisons}
\label{sec:AlgExp}
A set of tests were performed on the system to enhance non ellipse fitting algorithm. The effect of number of agents, the error thresholds and various other paramters were investigated. The result of these tests are shown in Figure \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}. Figure \ref{fig:swarmtesting} shows that as the number of agents increases the error value decrease (error of the segmentation algorithm). The final number of agents and maximum swarm iteration used in the system was based on a trade off between error achieved versus the computation time. Other algorithm parameters such as $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).

@@

 The main features affected by teh segmentation stage are the geometeric features computed in\textbf{FS1 }. Hence those are the features we used to estimate the efficency of the segmentation algorithm. The recognition accuracy using only the gometrical features is reported in Figure \ref{fig:testFeatonly} . The figure shows that \textsl{AlgS2} achieve better results.  %effect of the segmentation algorithm 

@@

\section{Conclusion and Future Work}
\label{ConclusionandFutureWork}
This paper presented a new approach to sketch recognition using PSO. The system uses both speed and curvature information which help improving the \textit{DPSO}. It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results. The use of statistical moments and structural features improves the recognition rate. The system was tested on 13 different symbols and achieved an overall accuracy of 93.5\%. The system does not depend on low level recognizer but rather on a set of high level features. This makes the system easily expandable to more and more symbols.

 A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition. Currently, the system processes one symbol at a time, an enhancement would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols. Another area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications.  
\bibliographystyle{IEEEbib}
\bibliography{../../neededfiles/Bibliographies/Mybibliography}
\end{document}

>>>>

\documentclass{article}
\usepackage{spconf}
\usepackage{amsmath}
\usepackage[dvips,pdftex]{graphicx}
\usepackage{subfigure}
\usepackage{named}
\usepackage[dvips,pdftex]{hyperref}
\title{Sketch Recognition using Particle Swarm Optimizations}
\twoauthors
  {Maha El Meseery, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
	{Signals Processing Group \\ Computers and Systems Department \\
  Electronic Research Institute\\Cairo, Egypt\\
	melmeseery@eri.sci.eg, samia@eri.sci.eg}
  {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
	{Computer Engineering Department\\
	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
	magdafayek@gmail.com,ndarwish@ieee.org}
\begin{document}
\maketitle
\begin{abstract}
Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. The results show that using PSO improves segmentation results which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.  
\end{abstract}

\section{Introduction}

Computer research nowadays aims at facilitating computer human interaction in every way. Pen-based interfaces give the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces. However, Using sketches as input to a computer system requires the system to understand the drawn sketches. In a truly freely sketch system this is not a trivial problem due to the sloppy and messy way users draw sketches. Sketches here stand for any hand drawing the user draws using digital stylus or similar device. Normally, sketches are composed of a set of strokes. The time between pen-up and pen-down events identify each stroke. The strokes are mainly the path of points, which the user path through between the pen-down and pen-up events while using the stylus or similar device.

Sketch recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}. The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods. Most stroke based systems either represent symbols as gestures or as a  hierarchical shape. Gestures are pen strokes drawn by the user that can immediately recognized by the system. A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,Ladder30,AlvaradoFreedom42,SketchRead2007}. Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them. Systems that employ such representation must first pre-process strokes to break them into geometric primitives. Recognition can then be treated as sub graph matching problem.

 In global shape properties approach, systems investigate general shape properties and its underlying strokes. Using shape properties (i.e. ratio of area of convex hull to area of shape) relaxes the assumption and role of each stroke in the symbol but it does not truly represent the appearance of the symbol\cite{DiagramOfflineConvexHull}. Unlike appearance based methods where individual strokes are disregarded and they focus on the overall appearance of shapes. In general, the  appearance based methods use image processing techniques to recognize sketches that the users draw.

 Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped. These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}. This let them to temporally group strokes and then recognizes them based on their strokes or global shape properties. \cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.

 Recognition based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}. The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process. Natural sketch are difficult to segment and group to create reliable fragment of geometric primitives. This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols. Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch.

In this paper we choose to implement a hybrid method between shape property and stroke based approaches. Our method tries to avoid many of the problems faced by approaches based on individual strokes or global shape properties while trying to gain the additional information that the stroke based system use. Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes. Users can draw symbols using any number of strokes in any order they like. The system segments each stroke using two different PSO algorithms. The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}. Our algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results. To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs. After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced. The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed.

 %The preprocessing phase captures user input stroke points and collects basic information about the stroke. In the segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments. In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier.

%This paper introduces a new sketch recognition system, which uses  % uses the computed feature vector to classify the segments into one of the previously trained classes.

 The remaining of the paper is as follows; the literature review on stroke segmentation is in Section \ref{sec:review}. Section \ref{sec:ParticleSwarmAlgorithm} introduces Particle Swarm Optimization. Section \ref{Sysdisc} describes the system in details.  Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively. Section \ref{sec:Experiments} describes the experiments that were preformed. We conclude in Section \ref{ConclusionandFutureWork}.

\section{Literature Review}
\label{sec:review}

Yu \cite{meanshift10} introduced a \textit{feature area} for each primitive and then computed the segmentation error for different types of primitives based on the computed \textit{feature area}. His system achieved good accuracy in simple shapes (square, ellipse,...etc) but did not perform well in complex shapes. A hybrid algorithm was introduced in \cite{earlyprocess} where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set. However, this system is limited to recognizing only specific simple geometric shapes with a set of low level recognizers. Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from input stroke.

A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. Chen et al. \cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system. Also, they did not take advantage of the curvature or local geometric properties of the digital curve. Yin \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts  Yin \cite{PolygonApproximationPSO} method and tries to improve it by adding curvature and other local information while segmenting strokes to achieve better segmentations. The next section presents the general particle swarm algorithm which is used in both our segmentation algorithms.

 % are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}

@@

After the system computes the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvature. However, using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of each of the four calculated information  is calculated. This mean is taken as the threshold \textit{th} which is used to separate the curve into regions ($R_i$) (Figure\ref{fig:timediff}); each region $R_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}. Those regions are further processed to find the maximum point $Max(R_i)$ of each region $R_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. The system repeats this process for curvature, time difference and speed information. All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate point $P_{pd}$ as the preprocessing. It is clear that some of the $P_{pd}$ points are redundant as specified in the preprocessing stage. % (as shown are redundant)
\begin{figure}
	\centering
		\subfigure[Possible dominate point]{ 	\label{fig:LabelsPPD}	\includegraphics[scale=0.7]{images/ppd.jpg}}
		\hfill
			\subfigure[DPSO algorithm encoding]{	\label{fig:LabelsPSO} \includegraphics[scale=0.7]{images/pso1.jpg}}			
	\caption{Segmentation}% a) Possible dominate point b) Particle encoding  
	\label{fig:pso1}
\end{figure}

\subsection{Segmentation}
\label{seg}
In the segmentation stage the input stroke is divided into a set of primitives. As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}. The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves. After two different segmentations are generated, the system chooses the segmentation with the minimum error. The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke. % The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing. The output is a set of dominant points which are connected by either lines or curves.

%///AllBlockDiagram.jpg  you can change the block to the main block. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.48]{images/blockSmall.jpg}
	\caption{Segmentation Block}
	\label{fig:segblock}
\end{figure}
\subsection{Ellipse fitting}
%The process starts by computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the ellipse center. The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.  
This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the center of the ellipse. The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box. The least square fitting algorithm \cite{chernov-2003} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})

@@

 where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke. A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.  After few iteration, the final fit error of the estimated ellipse is reported. Another measure is used to compute the efficiency of the final estimated ellipse. Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered. This eliminates fitting a  line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.

@@

 \subsubsection{\textit{Hybrid Fitting \textsl{AlgS2}}}
The algorithm has the same problem formulation but different fitness and error functions are used. An attempt is made to fit each segment into a line or circular arc. The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$. The approximation with the lower error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}. The sum of the approximation errors of all segments is the total error of the particle.  The total error of the particle is computed by equation %\ref{eq:errorSwarm2}
 \begin{equation}
E=\sum\nolimits_{i = 0}^M e(D_i) 
\label{eq:errorSwarm2}
\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.  The fitness is computed by the equation %\ref{eq:fitnessSwarm2} 
\begin{equation}
\max fitness(P_i ) = \frac{1}{{E \times M^k }}
\label{eq:fitnessSwarm2}
\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments. The larger $k$ is, the more effect the number of segments will have. For our system, $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.

@@

\section{Experimental}
\label{sec:Experiments}
The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system. The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. The dataset was divided into training set and test set. Four different splits for the training and test data are generated from the dataset. The results displayed are the average recognition accuracy of the four splits. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test. %In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.  
\begin{figure}
\centering 
\label{fig:symbolSet}
\fbox{ \parbox{5cm}{% 
		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
		\caption{The Symbol Set} Symbols used in the dataset.
\end{figure}
\subsection{Algorithms comparisons}
\label{sec:AlgExp}
A set of tests were performed on the system to enhance non ellipse fitting algorithm. The effect of number of agents, the error thresholds and various other parameters were investigated. The result of these tests are shown in Figure \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}. Figure \ref{fig:swarmtesting} shows that as the number of agents increases the error value decrease (error of the segmentation algorithm). The final number of agents and maximum swarm iteration used in the system was based on a trade off between error achieved versus the computation time. Other algorithm parameters such as $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).

@@

 The main features affected by the segmentation stage are the geometric features computed in\textbf{FS1 }. Hence those are the features we used to estimate the efficiency of the segmentation algorithm. The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} . The figure shows that \textsl{AlgS2} achieve better results.  %effect of the segmentation algorithm 

@@

\section{Conclusion and Future Work}
\label{ConclusionandFutureWork}
This paper presented a new approach to sketch recognition using PSO. The system uses both speed and curvature information which help improving the \textit{DPSO}. It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results. The final recognition is based on both global shape and hierarchical stroke properties. Adding stroke and segment geometrical features improved the final recognition. Global shape and moment properties was used to prevent the need to use graph matching which is computationally expensive. The system was tested on 13 different symbols and achieved an overall accuracy of 93.5\%. The system does not depend on low level recognizer but rather on a set of high level features. This makes the system easily expandable to more and more symbols.

 A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition. Currently, the system processes one symbol at a time, an enhancement would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols. Another area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications. Features that represents the appearance of the symbols are likely to improve the recognition that have similar geometrical structure (i.e. Nand and And gate in logic symbols).   
\bibliographystyle{IEEEbib}
\bibliography{../../neededfiles/Bibliographies/Mybibliography}
\end{document}
